On The Difficulty of Computing Non-Commutative Signature Gröbner Bases

### 2022-01-26 article_issac.tex
Reprise s -> ss dans l'exemple du langage non algébrique

If $\alpha \in \mathscr{F}_r$, it can be uniquely
written in the basis 
$\MM \left( \mathscr{F}_r \right) $
and the biggest monomial
involved is its signature
$\mathfrak{s}(\alpha) \in
\mathscr{F}_r$.
\fixmecc{c'est $\mathfrak{s}(\alpha) \in \MM (\mathscr{F}_r)$. Aussi définir
$\mathfrak{s}(E)$ pour un sous ensemble $E$.}

@@

Looking at principal syzygies and the algorithm for computing new ones, we wonder if an automaton can recognize leading monomials of syzygies in the case where the Gröbner basis is finite. Let us show that this is not the case using the non regular language\linebreak $\{a^nb^n \;|\; n \in \N\}$. Let us consider the alphabet 
$$\{a, b, c, d, e, s, t, x, y, z, \infty\}$$
and polynomials
\fixmecc{s -S ss partout}
\begin{align*}
g_1 & := sa - tcx \\
g_2 & := xa - ax \\
g_3 & := xd - dx \\
g_4 & := ya - ay \\
g_5 & := yd - dy \\
g_6 & := xbb - dyb \\
g_7 & := cya - ccx \\
g_8 & := xbe - dze \\
g_9 & := dz - zb \\
g_{10} & := cz - za \\
g_{11} & := s-tz
\end{align*}
together with the order $m_1 < m_2$ if:
\fixmecc{on compare d'abord par degrés}
\begin{itemize}
\item there are less $\infty$ symbols in $m_1$ than in $m_2$,
\item $m_1$ and $m_2$ have the same number of $\infty$ symbols and there are less $s$ symbols in $m_1$ than in $m_2$,
\item $m_1$ and $m_2$ have the same numbers of $\infty$ and $s$ symbols, but
$m_1 <_{deglex} m_2$ where\\ $a < b < c < d < x < y < z < e < t < s < \infty$.
\end{itemize}

\fixmecc{dire que cet ordre est fair}

\fixmecc{Given $l,r,l',r'\in\KX$, have (let?) $l\varepsilon_i r\leq l'\varepsilon_j r'$ if
  \begin{itemize}
  \item $l\LM(g_i)r < l'\LM(g_j)r'$
  \item $l\LM(g_i)r = l'\LM(g_j)r'$ et $i < j$
  \item $l\LM(g_i)r = l'\LM(g_j)r'$ et $i = j$ et $\deg{l} < \deg{l'}$.
  \end{itemize}
      {\bf dire que l'ordre est fair}

@@

For a better understanding, let us consider the case $n=2$. The monomial $\varepsilon_{11}a^2b^2e$ is a leading monomial of a syzygy because of the following sequence of equalities.
\begin{align*}
saabbe & =_{1} tcxabbe =_{2} tcaxbbe =_{6} tcadybe \\
& =_{5} tcaydbe =_{4} tcyadbe =_{7} tccxdbe \\
& =_{3} tccdxbe =_{8} tccddze =_{9} tccdzbe \\
& =_{9} tcczbbe =_{10} tczabbe =_{10} tzaabbe \\
& =_{11} saabbe
\end{align*}

@@

  Now, let us show that leading monomials of syzygies of the form $\varepsilon_{11}a^nb^me$ are such that $n = m$.
  Let $\sigma$ be such a syzygy and let $G$ be the graph whose vertices are elements of $\left\langle X \right\rangle$, and there is an edge between vertices of the form $lXr$ and $lYr$ as soon as $lg_ir$ occurs with a nonzero coefficient in $\sigma$ and $g_i=X - Y$. Let $W$ be the set of vertices of the following form:
  \begin{align*}
    & sa^nb^ne, \\
    & tc^pa^qxa^rd^nb^me, \\
    & tc^pa^qd^rxd^nb^me, \\
    & tc^pa^qya^rd^nb^me, \\
    & tc^pa^qd^ryd^nb^me, \\
    & tc^pza^qb^re, \\
    & tc^pd^qzb^re
  \end{align*}
  Each vertex in $W$ is related to at most two vertices, which also belong to $W$.
  Hence, the connected component of $sa^nb^me$ in $G$ is either a chain or a cycle (\fixmecc{écrire la phrase explicitement}).
  But since $\sigma$ is a syzygy, this connected component has to be a cycle, which implies $n = m$.

>>>>

If $\alpha \in \mathscr{F}_r$, it can be uniquely
written in the basis 
$\MM \left( \mathscr{F}_r \right) $
and the biggest monomial
involved is its signature
$\mathfrak{s}(\alpha) \in \MM (\mathscr{F}_r)$.
Moreover, given a subset $E$ of $\mathscr{F}_r$, we let
\[\mathfrak{s}(E)=\left\{\mathfrak{s}(\alpha)\mid \alpha\in E\right\}.\]

@@

Looking at principal syzygies and the algorithm for computing new ones, we wonder if an automaton can recognize leading monomials of syzygies in the case where the Gröbner basis is finite. Let us show that this is not the case using the non regular language\linebreak $\{a^nb^n \;|\; n \in \N\}$. Let us consider the alphabet 
$$\{a, b, c, d, e, s, t, x, y, z, \infty\}$$
and polynomials
\fixmecc{s -S ss partout}
\begin{align*}
g_1 & := ssa - tcx \\
g_2 & := xa - ax \\
g_3 & := xd - dx \\
g_4 & := ya - ay \\
g_5 & := yd - dy \\
g_6 & := xbb - dyb \\
g_7 & := cya - ccx \\
g_8 & := xbe - dze \\
g_9 & := dz - zb \\
g_{10} & := cz - za \\
g_{11} & := ss-tz
\end{align*}
together with the order $m_1 < m_2$ if:
\begin{itemize}
\item the degree of $m_1$ is strictly smaller than the degree of $m_2$,
\item $m_1$ and $m_2$ have the same degree and there are less $\infty$ symbols in $m_1$ than in $m_2$,
\item $m_1$ and $m_2$ have the same degree, the same number of $\infty$ symbols and there are less $s$ symbols in $m_1$ than in $m_2$,
\item $m_1$ and $m_2$ have the same degree, the same numbers of $\infty$, and the same number $s$ symbols, but
$m_1 <_{lex} m_2$ where\\ $a < b < c < d < x < y < z < e < t < s < \infty$.
\end{itemize}
Such defined, this order is fair.

\fixmecc{Given $l,r,l',r'\in\KX$, have (let?) $l\varepsilon_i r\leq l'\varepsilon_j r'$ if
  \begin{itemize}
  \item $l\LM(g_i)r < l'\LM(g_j)r'$
  \item $l\LM(g_i)r = l'\LM(g_j)r'$ et $i < j$
  \item $l\LM(g_i)r = l'\LM(g_j)r'$ et $i = j$ et $\deg{l} < \deg{l'}$.
  \end{itemize}
  Such defined this order is fair.

@@

For a better understanding, let us consider the case $n=2$. The monomial $\varepsilon_{11}a^2b^2e$ is a leading monomial of a syzygy because of the following sequence of equalities.
\begin{align*}
ssaabbe & =_{1} tcxabbe =_{2} tcaxbbe =_{6} tcadybe \\
& =_{5} tcaydbe =_{4} tcyadbe =_{7} tccxdbe \\
& =_{3} tccdxbe =_{8} tccddze =_{9} tccdzbe \\
& =_{9} tcczbbe =_{10} tczabbe =_{10} tzaabbe \\
& =_{11} ssaabbe
\end{align*}

@@

  Now, let us show that leading monomials of syzygies of the form $\varepsilon_{11}a^nb^me$ are such that $n = m$.
  Let $\sigma$ be such a syzygy and let $G$ be the graph whose vertices are elements of $\left\langle X \right\rangle$, and there is an edge between vertices of the form $lXr$ and $lYr$ as soon as $lg_ir$ occurs with a nonzero coefficient in $\sigma$ and $g_i=X - Y$. Let $W$ be the set of vertices of the following form:
  \begin{align*}
    & ssa^nb^ne, \\
    & tc^pa^qxa^rd^nb^me, \\
    & tc^pa^qd^rxd^nb^me, \\
    & tc^pa^qya^rd^nb^me, \\
    & tc^pa^qd^ryd^nb^me, \\
    & tc^pza^qb^re, \\
    & tc^pd^qzb^re
  \end{align*}
  Each vertex in $W$ is related to at most two vertices, which also belong to $W$. When an undirected graph is such that each vertex is related to at most two vertices, each connected component is either a chain or cycle, so that the connected component of $ssa^nb^me$ in $G$ is either a chain or a cycle.
  But since $\sigma$ is a syzygy, this connected component has to be a cycle, which implies $n = m$.
